Staged Abstract Interpreters

### 2018-10-29 paper/10_intro.tex
intro

Recent research uncovers a systematic approach to construct sound abstract interpreters 
from concrete interpreters \cite{DBLP:journals/jfp/HornM12, DBLP:conf/icfp/HornM10, DBLP:journals/pacmpl/DaraisLNH17}.
Given the structural similarity between concrete and abstract interpreters, intellectually 
it is natural to wonder what would be a staged abstract interpreter, and how to effectively 
specializing an abstract interpreter having considered their functional dissimilarity. 
This paper studies the application of the first Futamura projection on abstract interpreters, 
and the approach to construct efficient abstract interpreters by multi-stage programming.

@@

The slogan of multi-stage programming is "abstraction without regret". 
Of course, all interpreters are metalinguistic abstractions, but some interpreters are more abstract 
than others \todo{maybe rephrase}. 
One technical challenge is to deal with the non-determinisism, fixed-point and abstract domains in 
abstract interpreters when staging is introduced. The abstract interpreter, as a semantic artifact, 
should be written in a style that is easy for people to communicate the formulation, but also be friendly
to specialization and generating efficient code. \todo{We discuss the form of abstract interpreters for staging}
We show the use LMS framework, that enables multi-stage programming by just changing the types 
of the analyzer program, and we thus bridge the gap between designing sound program analyses 
and implementing efficient program analyzers.

@@

The use of staged abstract interpreters on open programs improves the scalability.
One challenge in static analysis of modern software is that the programs are usually 
shipped with large library code, for example, it is observed that analyzing a simple
"Hello World" program in Java depends on additional 3,000 classes in the library \cite{DBLP:conf/oopsla/KulkarniMZN16}.
Soundly and precisely analyzing these libraries sometimes are expensive and unnecessary repeatedly.
When applying the staged abstract interpreter on libraries, we leave the unknown 
calling context as dynamic value and generate partial analyzing result which is 
repsented as a residul program. The partial analyzing result can be resued and composed
with the analyzing result of application code later. \todo{we show this can be applied 
to context/path/flow-sensitive even with different sensitivity parameters.}
It has been observed library code requires less precision \todo{cite}.

% Many optimization of static analysis are ad-hoc.

@@

\begin{itemize}
\item Starting from a definitional abstract interpreter for a higher-order functional language,
  which the syntax and semantics are described in section \ref{bg_lang},
  we show how to turn it into a staged abstract interpreter by adding stage annotations within
  Lightweight Modular Staging (LMS) framework, a library-based approach to code generation. 
  We later extend the staged abstract interpreter to support imperative language features 
  (section \ref{cases_imp}), such as states, loops and exceptions.
\item We elaborate the idea that the staged abstract interpreter not only improves the \textit{performance}
  but also the \textit{scalability}, which are two major issues in static analysis, if we apply it on open
  programs. \todo{more on this point.}
\item To evaluate the generality and applicability of the idea, we conduct three case studies in section \ref{cases_study}. 
  First we revisited Abstract Compilation \cite{Boucher:1996:ACN:647473.727587}, which is a closure generation 
  technique applied on 0-CFA. We show that it can be understood and easily implemented as an instance of 
  staged abstract interpreter. 
  Second, we demonstrate sophisticated control-flow analysis techniques, such as adaptive context/flow/path-sensitive 
  analysis and store-widening can be integrated with staged abstract interpreters without losing any precision.
  Finally, we show that in an imperative language, the idea can be applied onto an abstract interpreter with
  an abstract domain that performs numerical analysis.
  (Optional: taint/information flow analysis?)
\item We also empirically evaluate the performance improved by staging an big-step abstract interpreter 
  for control-flow analysis. We compare against with the unstaged version, a small-step version (unstaged), 
  and an optimizing small-step version.
\end{itemize}

>>>>

Recent research uncovers a systematic approach to construct sound abstract interpreters 
from concrete interpreters \cite{DBLP:journals/jfp/HornM12, DBLP:conf/icfp/HornM10, DBLP:journals/pacmpl/DaraisLNH17}.
Given the structural similarity between concrete and abstract interpreters, intellectually 
it is natural to raise the question whether it is possible to stage an abstract interpreter, and how 
to effectively specialize an abstract interpreter having considered their functional dissimilarity. 
This paper studies the application of the first Futamura projection on abstract interpreters, 
and the approach to construct efficient abstract interpreters by multi-stage programming.

@@

The slogan of multi-stage programming is "abstraction without regret". 
Of course, all interpreters are metalinguistic abstractions, but some interpreters are more abstract 
than others \todo{maybe rephrase}. 
One technical challenge is to deal with the non-determinisism, fixed-point iterations and abstract 
domains in abstract interpreters when staging is introduced. The abstract interpreter, as a semantic artifact, 
should be written in a style that is easy for people to communicate the formulation, but also be friendly
to specialization and generating efficient code. \todo{We discuss the form of abstract interpreters for staging}
We show the use LMS framework, that enables multi-stage programming by just changing the types 
of the analyzer program, and we thus bridge the gap between designing sound program analyses 
and implementing efficient program analyzers.

@@

The use of staged abstract interpreters on open programs improves the scalability.
One challenge in static analysis of modern software is that the programs are usually 
shipped with large library code, for example, it has been shown that analyzing a simple
"Hello World" program in Java depends on additional 3,000 classes in the library \cite{DBLP:conf/oopsla/KulkarniMZN16}.
Soundly and precisely analyzing these libraries sometimes are expensive and unnecessary repeatedly.
When applying the staged abstract interpreter on libraries, we leave the unknown 
calling context as dynamic value and generate partial analyzing result which is 
repsented as a residul program. The partial analyzing result can be resued and composed
with the analyzing result of application code when available later. 
It has been observed that partially applying context-sensitivity on selected portion of the program could
improve the precision and efficiency \cite{zipper2018, Kastrinis:2013:HCP:2491956.2462191}. 
We show that staging abstract interpreters as an approach to effectively implement hybird context-sensitivity\todo{}.

staged polymorphism, decide when to compile an analysis.  
% Many optimization of static analysis are ad-hoc.

@@

\begin{itemize}
\item Starting from a definitional abstract interpreter for a higher-order functional language,
  which the syntax and semantics are described in section \ref{bg_lang},
  we show how to turn it into a staged abstract interpreter by adding stage annotations within
  Lightweight Modular Staging (LMS) framework, a library-based approach to code generation. 
  We later extend the staged abstract interpreter to support imperative language features 
  (section \ref{cases_imp}), such as states, loops and exceptions.
\item We elaborate the idea that the staged abstract interpreter not only improves the \textit{performance}
  but also the \textit{scalability}, which are two major issues in static analysis, if we apply it on open
  programs. \todo{more on this point.}
\item To evaluate the generality and applicability of the idea, we conduct three case studies in section \ref{cases_study}. 
  First we revisited Abstract Compilation \cite{Boucher:1996:ACN:647473.727587}, which is a closure generation 
  technique applied on 0-CFA. We show that it can be understood and easily implemented as an instance of 
  staged abstract interpreter. 
  Second, we demonstrate sophisticated control-flow analysis techniques, such as adaptive context/flow/path-sensitive 
  analysis and store-widening can be integrated with staged abstract interpreters without losing any precision.
  Finally, we show that in an imperative language, the idea can be applied onto an abstract interpreter with
  an abstract domain that performs numerical analysis.
  (Optional: taint/information flow analysis?)
\item We also empirically evaluate the speed improved by staging an big-step abstract interpreter 
  for control-flow analysis. We compare against with the unstaged version, a small-step version (unstaged), 
  and an optimizing small-step version.
\end{itemize}

### 2018-10-29 paper/70_related.tex
intro
continue working on outline
misc

@@

\textbf{Abstracting Abstract Machine}. Abstracting Definitional Interpreter \cite{darais2017abstracting}

@@

>>>>

Abstract compilation for analyzing logic programs\cite{DBLP:conf/flops/AmatoS01}

@@

\textbf{Abstract Interpreters}.

Abstract interpretation \cite{DBLP:conf/popl/CousotC77}.

Abstracting Definitional Interpreter \cite{darais2017abstracting}

Polyhedra, online decomposition

\textbf{Modular Static Analysis}.

Cousot \& Cousot \cite{DBLP:conf/cc/CousotC02}, 
Compositional shape analysis \cite{DBLP:conf/popl/CalcagnoDOY09}

\textbf{Multi-Stage Programming}.

@@

TODO: compare to other staging approach (e.g., MetaOCaml)

### 2018-10-29 paper/abstract.tex
intro

We also empirically evaluate the improvement of performance on benchmark programs for a subset of Scheme.
The result shows ...

>>>>

We also empirically evaluate the improvement of performance on control-flow analysis of benchmark programs.
The result shows an average speedup of Nx when staging to Scala for a monovariant analysis, and Mx for polyvariant analysis.
